---
title: "Project 2"
author: "Santiago Torres"
date: "10/03/2021"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: vignette
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jsonlite)
library(dplyr)
library(tidyverse)
library(ggplot2)
```

# League of Legends Dataset 1


## Data Storage

### Read in JSON file


Here we read in the data from the League of Legends [website](https://ddragon.leagueoflegends.com/cdn/11.19.1/data/en_US/champion.json) and create a tibble from the JSON for easier use within R
```{r}
champ_data <- fromJSON("https://ddragon.leagueoflegends.com/cdn/11.19.1/data/en_US/champion.json")

# 
champ_df <- tribble(~name,~attack,~defense,~difficulty,~tags,
                    "Starter", 0, 0, 0, "Fighter, Mage")
```

### Convert JSON to Dataframe

Iterate through the JSON file and pull the columns into the tibble we created.

```{r}

for (champ in champ_data$data) { 
  # names
 champ_df <- champ_df %>% add_row(name = champ[4]$name, attack = champ[[7]][1]$attack, defense = champ[[7]][2]$defense, difficulty = champ[[7]][4]$difficulty, tags =  paste(champ[9]$tags, collapse = ", "))
}

#remove dummy row:
champ_df <- champ_df[2:nrow(champ_df),]
```



### Write to CSV

```{r}

output_filename <- str_c(getwd(), "/champions.csv")

write.csv(champ_df,output_filename)

```


## Importing and Preparing Data

Let's take a look at our data 

```{r}

champ <- read.csv(output_filename)

head(champ)

```
### Set Row Names

```{r}
row.names(champ) <- champ$name
```

This is already pretty clean, but we should normalize some of the `tags` data so that we can do more analysis later on. 

I'm going to keep things simple for now and simply take the first tag as the canonical `primary_role` and the second as the `secondary_role`

```{r}
primary_role <- c()
secondary_role <- c()
for (tag in champ$tags) {
  roles <- tag %>% str_split(",")
  primary_role <- c(primary_role, roles[[1]][1] %>% trimws())
  secondary_role <- c(secondary_role, roles[[1]][length(roles[[1]])] %>% trimws())
}

champ <- champ %>% add_column(primary_role = primary_role, secondary_role = secondary_role)

head(champ)

# drop extra column

champ <- champ[,2:length(champ)]
```


This allows us to perform more data analysis later.

Let's also create a new column called `strength` so we can compare the summation of a champions attack and defense.

```{r}
champ <- champ %>% mutate(strength = attack + defense)
```


## Data Analysis

We can do some some high level sorting to get a sense of the 5 weakest champions and the 5 strongest champions by each category

### Strength by primary_role

```{r}

champ %>% group_by(primary_role) %>% summarize(max = max(strength), min = min(strength), mean = mean(strength), stdev = sd(strength), count = n())

```


### Attack by primary_role

```{r}

champ %>% group_by(primary_role) %>% summarize(max = max(attack), min = min(attack), mean = mean(attack), stdev = sd(attack), count = n())

```
I'm surprise that Supports come out higher than Mages on attack as Mages should in theory have more attacking power than a Support class champion.  
### Defense by primary_role

```{r}

champ %>% group_by(primary_role) %>% summarize(max = max(defense), min = min(defense), mean = mean(defense), stdev = sd(defense), count = n())

```
Understandly, Tanks come out with the highest defense.

We can add a third dimension, `difficulty` by adding in a scatter plot to get a graphical sense of the data.

```{r}
champ %>% ggplot(aes(x=difficulty, y=strength, color = primary_role)) + geom_point() 
```
It's hard to tell if strength increases as difficulty increases, so we can also split out the graphs by the role to get a better sense of it


```{r}
champ %>% filter(primary_role == "Fighter") %>% ggplot(aes(x=difficulty, y=strength, color = primary_role)) + geom_point() + geom_smooth(method=lm) + facet_wrap(~primary_role)
```



# Child Mortality Rates Dataset 2

```{r}


output_filename <- str_c(getwd(), "/champions.csv")

write.csv(champ_df,output_filename)
```

## Importing and Preparing Data


```{r}

```


## Data Analysis

## Importing and Preparing Data