---
title: "Project 2"
author: "Santiago Torres"
date: "10/03/2021"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: vignette
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jsonlite)
library(dplyr)
library(tidyverse)
library(ggplot2)
```

# League of Legends Dataset 1


## Data Storage

### Read in JSON file


Here we read in the data from the League of Legends [website](https://ddragon.leagueoflegends.com/cdn/11.19.1/data/en_US/champion.json) and create a tibble from the JSON for easier use within R
```{r}
champ_data <- fromJSON("https://ddragon.leagueoflegends.com/cdn/11.19.1/data/en_US/champion.json")

# 
champ_df <- tribble(~name,~attack,~defense,~difficulty,~tags,
                    "Starter", 0, 0, 0, "Fighter, Mage")
```

### Convert JSON to Dataframe

Iterate through the JSON file and pull the columns into the tibble we created.

```{r}

for (champ in champ_data$data) { 
  # names
 champ_df <- champ_df %>% add_row(name = champ[4]$name, attack = champ[[7]][1]$attack, defense = champ[[7]][2]$defense, difficulty = champ[[7]][4]$difficulty, tags =  paste(champ[9]$tags, collapse = ", "))
}

#remove dummy row:
champ_df <- champ_df[2:nrow(champ_df),]
```



### Write to CSV

```{r}

output_filename <- str_c(getwd(), "/champions.csv")

write.csv(champ_df,output_filename)

```


## Importing and Preparing Data

Let's take a look at our data 

```{r}

champ <- read.csv(output_filename)

head(champ)

```
### Set Row Names

```{r}
row.names(champ) <- champ$name
```


### Tidy and Prep Data

This is already pretty clean, but we should normalize some of the `tags` data so that we can do more analysis later on. 

I'm going to keep things simple for now and simply take the first tag as the canonical `primary_role` and the second as the `secondary_role`

```{r}
primary_role <- c()
secondary_role <- c()
for (tag in champ$tags) {
  roles <- tag %>% str_split(",")
  primary_role <- c(primary_role, roles[[1]][1] %>% trimws())
  secondary_role <- c(secondary_role, roles[[1]][length(roles[[1]])] %>% trimws())
}

champ <- champ %>% add_column(primary_role = primary_role, secondary_role = secondary_role)

head(champ)

# drop extra column

champ <- champ[,2:length(champ)]
```


This allows us to perform more data analysis later.

Let's also create a new column called `strength` so we can compare the summation of a champions attack and defense.

```{r}
champ <- champ %>% mutate(strength = attack + defense)
```


## Data Analysis

We can do some some high level sorting to get a sense of the strength of each role

### Strength by primary_role

```{r}

champ %>% group_by(primary_role) %>% summarize(max = max(strength), min = min(strength), mean = mean(strength), stdev = sd(strength), count = n())

```

It seems like the strongest groups are Fighters and Tanks, with Fighter's slightly edging out to due a higher mean and a lower spread.


### Attack by primary_role

```{r}

champ %>% group_by(primary_role) %>% summarize(max = max(attack), min = min(attack), mean = mean(attack), stdev = sd(attack), count = n())

```


I'm surprised that Supports come out higher than Mages on attack as Mages should in theory have more attacking power than a Support class champion.  


### Defense by primary_role

```{r}

champ %>% group_by(primary_role) %>% summarize(max = max(defense), min = min(defense), mean = mean(defense), stdev = sd(defense), count = n())

```
Understandly, Tanks come out with the highest defense.

We can add a third dimension, `difficulty` by adding in a scatter plot to get a graphical sense of the data.
### Difficulty by Strength

```{r}
champ %>% ggplot(aes(x=difficulty, y=strength)) + geom_point() + geom_smooth(method=lm)
```

It looks like there is a slightly negative slope with strength as difficulty decreases. However, remembering Simpson's Paradox, we can stratify our data into the primary roles to see what happens when we evaluate it by role.

```{r}
champ %>% ggplot(aes(x=difficulty, y=strength, color = primary_role)) + geom_point() + geom_smooth(method=lm) + facet_wrap(~primary_role)
```

You can see that Tanks increase in strength drastically as difficulty increases and Assassins actually drop drastically as difficulty increases.


## Overview and Conclusion

The assigned problem was to find the highest starting HP for each champion, but after taking a look at the data, I felt that a more insightful analysis would be to look at the relative strength of each champion role by its primary role. My key assumptions were as follows:

  + Strength is an additive representation of attack and defense (Attack + Defense).
  + A champion's primary role in the data is the first in the list of roles provided.
  
Based off of those assumptions, the strongest groups in League of Legends are Fighters and Tanks. 

We could do more statistical analysis to determine the exact correlation between strength and difficulty as well as more targeted analysis with the attack / defense characteristics.


# Child Mortality Rates Dataset 2

## Overview of Approach

Alec wants to answer the following questions with this dataset:

  + What 10 countries have the highest under-5 mortality rates today?
  + For the 10 worst countries, visualize the under-5 mortality trend over time.
  + Comparatively, what does the trend for G7 countries look like over time?

## Data Storage

I downloaded the data from [The World Bank](https://data.worldbank.org/indicator/SH.DYN.MORT) and uploaded the csv file to my GitHub account with a wide format (years as columns). The data is about Mortality rate, under-5 (per 1,000 live births)


## Importing and Preparing Data
```{r}
input_csv <- "https://raw.githubusercontent.com/sserrot/DATA607/main/Project%202/child_mortality_rates/API_SH.DYN.MORT_DS2_en_csv_v2_3012069.csv"
child_mort <- read.csv(input_csv, check.names = FALSE)
colnames(child_mort)
```

### Renaming Column Headers

```{r}
child_mort <- child_mort %>% rename(country_name = "﻿Country Name", country_code = "Country Code")

head(child_mort)
```
### Lengthen the data

```{r}
child_mort <- child_mort %>% pivot_longer(cols = !c("country_name","country_code"),names_to = "year", names_transform = list(year = as.integer), values_to =  "child_mortality")

head(child_mort)
```

Aruba has no data so we can just look at the US to make sure we brought in the data as expected.

```{r}
child_mort %>% filter(country_code == "USA")
```



## Data Analysis

### Question 1: What 10 countries have the highest under-5 mortality rates today?

Let's first find the most recent year that we have data for:

```{r}
 max(child_mort$year)
```

Let's filter the data for 2019 and slice the first 5 countries with the highest child mortality 

```{r}
child_mort %>% filter(year == 2019) %>% arrange(desc(child_mortality)) %>% slice(1:10)
```

Let's store this in a list so we can reference these countries later.

```{r}
bottom_ten_code <- child_mort %>% filter(year == 2019) %>% arrange(desc(child_mortality)) %>% slice(1:10) %>% select(country_code)

```

### Question 2: For the 10 worst countries, visualize the under-5 mortality trend over time.

Since we have the list of the 10 countries with the worst child mortality, we can filter our original dataset and plot the trend of the data over time.

#### Filter

```{r}
child_mort_bottom_ten <- child_mort[child_mort$country_code %in% bottom_ten_code$country_code,]
```


#### Plot countries over time

```{r}
child_mort_bottom_ten %>% ggplot(aes(x=year,y=child_mortality, color = country_name)) + geom_point() 
```

It looks like all these countries have dropped their child mortality rates by between 1/2 to 1/4 since the data has been tracked. We can get the exact numbers from the following snippet of code:

```{r}

earliest_year <- child_mort_bottom_ten %>% drop_na() %>% group_by(country_name) %>% summarize(min_year = min(year))

```


### Question 3: Comparatively, what does the trend for G7 countries look like over time?

First, we have to find the list of G7 countries from [Wikipedia](https://en.wikipedia.org/wiki/Group_of_Seven)

Canada, France, Germany, Italy, Japan, the United Kingdom, and the United States:

```{r}
g_seven <- c("Canada", "France", "Germany", "Italy", "Japan","United Kingdom","United States")
```
#### Plot G7 over time

```{r}
child_mort %>% filter(country_name %in% g_seven) %>% ggplot(aes(x=year,y=child_mortality, color = country_name)) + geom_point() 
```

The G7 countries have also dropped their mortality rates by about 1/3 to 1/5, although their starting points are much lower relatively speaking.


# Tuition Assistance Program (TAP) Fall Headcount by College - 2020

## Overview of Approach

Bianka wants to answer the following questions with this dataset:

  + See recipient trends by academic year, by section type
  + Get percentage of students receiving TAP by sector type or academic year.

## Data Storage

I downloaded the data from [data.gov](https://catalog.data.gov/dataset/tuition-assistance-program-tap-fall-headcount-by-college-sector-group-and-level-of-study-b) and uploaded the csv file to my GitHub account. The data is about all the trees in NYC in 2015.

## Importing and Preparing Data

```{r}
input_csv <- "https://raw.githubusercontent.com/sserrot/DATA607/main/Project%202/tuition_assistance/Tuition_Assistance_Program__TAP__Fall_Headcount_By_College__Sector_Group__and_Level_of_Study___Beginning_2000.csv"
TAP  <- read.csv(input_csv, check.names = FALSE)
colnames(TAP)
```
### Rename headers

```{r}
TAP <- TAP %>% rename(academic_year = "Academic Year", tap_college_code = "TAP College Code", fed_school_code = "Federal School Code", level = Level, study_level = "TAP Level of Study", sector_type = "Sector Type", sector_group = "TAP Sector Group", headcount = "TAP Fall Headcount", tap_college_name = "TAP College Name")

head(TAP)
```

This data has already been converted to a long format, so there is not much more tidying to do.



## Data Analysis

### Question 1 See recipient trends by academic year, by sector type


Let's start by looking through the data

```{r}
unique(TAP$academic_year)
```

We have data for the past 19 years, let's see the max headcount recipient per year

```{r}
TAP %>% group_by(academic_year, tap_college_name, sector_type) %>% summarize(largest_headcount_by_year = max(headcount)) %>% arrange(desc(largest_headcount_by_year))
```

Largest headcount recipient ever was `CUNY MANHATTAN CC` in 2015.


how many colleges are there?

```{r}
length(unique(TAP$tap_college_name))
```

That's a lot of colleges to represent in a single graph, so we should try to figure out a way to split out the dataset into a more manageable set of graphs.

```{r}
TAP %>% ggplot(aes(x=academic_year, y=headcount)) + geom_point() + facet_wrap(~sector_group)
TAP %>% ggplot(aes(x=academic_year, y=headcount)) + geom_point() + facet_wrap(~sector_type)
```

Still hard to tell what is going on. Maybe if we sum each group's total headcount per year

```{r}
TAP_sector_group <- TAP %>% group_by(academic_year, sector_group) %>% summarize(headcount = sum(headcount))

TAP_sector_group %>% ggplot(aes(x=academic_year, y=headcount, color = sector_group)) + geom_point() + facet_wrap(~sector_group)

TAP_sector_type <- TAP %>% group_by(academic_year, sector_type) %>% summarize(headcount = sum(headcount))

TAP_sector_type %>% ggplot(aes(x=academic_year, y=headcount, color = sector_type)) + geom_point()

```


Now we are getting somewhere, you can clearly see a drop in Private headcount as time progresses as well as Independent sector_group


### Question 2 Get percentage of students receiving TAP by sector type or academic year.

First we need to get the total number of students per academic year:

```{r}
TAP_academic_year <- TAP %>% group_by(academic_year) %>% summarize(total = sum(headcount))
```

Then we enrich the prior data with the new total per year:

```{r}
TAP_sector_group <- TAP_sector_group %>% left_join(TAP_academic_year)

TAP_sector_group <- TAP_sector_group %>% mutate(percent_students = format(headcount / total * 100, digits = 3))

TAP_sector_group %>% filter(percent_students == max(TAP_sector_group$percent_students))


TAP_sector_type <- TAP_sector_type %>% left_join(TAP_academic_year)

TAP_sector_type <-TAP_sector_type %>% mutate(percent_students = format(headcount / total * 100, digits = 3))

TAP_sector_type %>% filter(percent_students == max(TAP_sector_type$percent_students))
```


Turns out GGPlot can just calculate it for you when you graph a bar_plot with position set to `fill`, but it was still nice to see the max values for each group.


```{r}
TAP_sector_group %>% ggplot(aes(fill=sector_group,y=headcount,x=academic_year)) + geom_bar(position="fill", stat="identity")


TAP_sector_type %>% ggplot(aes(fill=sector_type,y=headcount,x=academic_year)) + geom_bar(position="fill", stat="identity")

```




